---
title: "R Fundamentals"
output: html_notebook
---
THe white space is for notes - markdown space
THe gray space - r chunk - coding where we write r codes and execute using ctrl+shift+enter

to add new r chunk - you can use ctrl+alt+i or else you can add it from above insert button on top

 

```{r}
```


```{r}
print("Hello World")
```


```{r}
x = 5
y = 6
print(x+y)
```

1. Data Types
  - primitive data types - integer, numeric,character, logical, complex
  - non primitive data types - vector, matrix, dataframe, list, arrays, factors
  
2. Control Flow
  - if else
  - for loop
  - while loop

3. Functions
  - built in functions
  - user defined functions
  
4. Packages and libraries

## Data Types

```{r}
# numeric
x = 5
y <- 6.6
10 -> z

print(x)
print(y)
print(z)
class(x)
class(y)
```

```{r}
# integer
m = 25L
print(m)
class(m)
```

```{r}
# character
k = "Hello world"
print(k)
class(k)
```
```{r}
# logical
k = TRUE
print(k)
class(k)
```
```{r}
# complex
m = 5 + 6i
print(m)
class(m)
```
## non primitive data types
### Vectors
  - used to store multiple values, all the values should be of same data type 
  - one dimensional
  - used to store data, mathematical compuations
  - can be defined using c()
```{r}
x = c(4,5,6,8)
print(x)
class(x)
```
```{r}
length(x)
```
```{r}
# add a new value to a vector
x = c(7,8,5,2)
print(x)
x = c(x,12)
print(x)
```
```{r}
#indexing
x[1]
x[3]
```
```{r}
# deleting a value based on index
x = x[-3]
print(x)
```
## Matrix
- an extension of vectors, it is of multiple dimensions
- all the values should be of one data type
- can be defined using - matrix()
```{r}
m = matrix(c(7,8,5,9,5,3,3,6),nrow = 2,ncol = 4)
print(m)
class(m)
dim(m)
```
```{r}
m[2,4]
```
```{r}
colnames(m) = c("Day1",'Day2','Day3','Day4')
m
```
```{r}
Day5 = c(12,15)
cbind(m,Day5)
```
```{r}
rbind(m,c(7,12,45,56))
```

### Dataframes
- used to handle 2D data, one column can of different data type than other column
- can be defined using data.frame()

```{r}
m = data.frame(
  Name = c("John",'Max','Ketty','Georgia','Marco'),
  Age = c(25,45,35,34,24),
  City = c("KUL","CGK","DXB","KUL","AUH")
)
class(m)
```
```{r}
m
```
```{r}
str(m)
```
```{r}
head(m)
tail(m)
colnames(m)
nrow(m)
dim(m)
```


```{r}
summary(m)
```
```{r}
m$Name
```
```{r}
df = read.csv("D:\\AI\\data\\datasets-1\\Bank_churn_modelling.csv")
dim(df)
```
```{r}
head(df)
```

```{r}
df$RowNumber = as.numeric(df$RowNumber)
head(df)
```
```{r}

```

