---
title: "R Notebook"
output: html_notebook
---

1. Domain Exploration, problem formulation
  - understand the business process, common beleifs, recent events
  
2. Data Collection, Data Exploration
  - collect data from multiple units, compile, for a dataset
  - perform a generic exploration to the dataset, identify the common challenges in the dataset
  
3. Data Cleaning
  - Handle duplicate entries
  - Handle missing values
  - Handle unwanted columns
  - Handle outliers
  
4. Exploratory Data Analytics / Root Cause Analytics
  - Data Visualization
    - Univariate Analysis
    - Bivariate Analsis
    - Multivariate Analysis
    
  - Statistical Research
    - Correlation Analysis
    - ANOVA, Chi square test
    
5. Prepare a report
  - Data/statistical Observations
  - Business inference of observations
  - Recommendations

## 2. Data Collection and Data Exploration

```{r}
#load libraries

```

```{r}
# load dataset
df = read.csv("D:\\AI\\data\\datasets-1\\Bank_churn_modelling.csv")
dim(df)
```
```{r}
head(df)
```
```{r}
str(df)
```
```{r}
unique(df$Gender)
unique(df$Geography)
```
```{r}
summary(df)
```


Observations - 
  - RowNumber, Customerid and surname are unwanted columns
  - approx more than 25% of the customers have 0 balance
  - age,balance may have outliers

## 3. Data Cleaning
