---
title: "R Notebook"
output: html_notebook
---

1. Domain Exploration, problem formulation
  - understand the business process, common beleifs, recent events
  
2. Data Collection, Data Exploration
  - collect data from multiple units, compile, for a dataset
  - perform a generic exploration to the dataset, identify the common challenges in the dataset
  
3. Data Cleaning
  - Handle duplicate entries
  - Handle missing values
  - Handle unwanted columns
  - Handle outliers
  
4. Exploratory Data Analytics / Root Cause Analytics
  - Data Visualization
    - Univariate Analysis
    - Bivariate Analsis
    - Multivariate Analysis
    
  - Statistical Research
    - Correlation Analysis
    - ANOVA, Chi square test
    
5. Prepare a report
  - Data/statistical Observations
  - Business inference of observations
  - Recommendations

## 2. Data Collection and Data Exploration

```{r}
#load libraries

```

```{r}
# load dataset
df = read.csv("D:\\AI\\data\\datasets-1\\Bank_churn_modelling.csv")
dim(df)
```
```{r}
head(df)
```
```{r}
str(df)
```
```{r}
unique(df$Gender)
unique(df$Geography)
```
```{r}
summary(df)
```
```{r}
length(unique(df$CustomerId))
```


Observations - 
  - RowNumber, Customerid and surname are unwanted columns
  - approx more than 25% of the customers have 0 balance
  - age,balance may have outliers

## 3. Data Cleaning
  - Handle duplicate entries
  - Handle missing values
  - Handle unwanted columns
  - Handle outliers
```{r}
# check for duplicates
sum(duplicated(df))
sum(duplicated(df$CustomerId))
```
THere are no duplicated rows presents
```{r}
# to see which rows are duplicated
df[duplicated(df),]
```

```{r}
# check for missing values
colSums(is.na(df))
```
```{r}
df[(is.na(df$Tenure)|(is.na(df$EstimatedSalary))),]
```
checking skewness
```{r}
library(moments)
skewness(df[,c(8,13)],na.rm=TRUE)
```
As the attributes tenure and salary are almost normally distributed, we will impute the missing values
by mean of the respective attribute
```{r}
df[is.na(df$Tenure),8] = mean(df$Tenure,na.rm=TRUE)
df[is.na(df$EstimatedSalary),13] = mean(df$EstimatedSalary,na.rm=TRUE)
colSums(is.na(df))
```
```{r}
colnames(df)
```
```{r}
# to drop all rows with missing values
na.omit(object = df)
# drop the rows where more than 50% of values are missing in the row
thresh = 0.5*ncol(df)

# keeping the rows where count of missing values are less than thresh
df = df[rowSums(is.na(df))<thresh,]
```

```{r}
# dropping the unwanted columns - rownumber, surname and customerID
df = df[,c(-1,-2,-3)]
```

```{r}
dim(df)
head(df)

```
```{r}
# check for outliers - using skewness approach
skewness(df[,c(-2,-3)])
```
## 4. Exploratory Data Analytics - Root Cause Analytics

  - Data Visualization
    - Univariate Analysis - studying and analysing individual variables
    - Bivariate Analsis - analsing effect of each variable on the target variable
    - Multivariate Analysis - analyse effect of multiple columns together on target variable
    
  - Statistical Research
    - Correlation Analysis
    - ANOVA, Chi square test
    
### Univariate Analysis
  - Categorical - discrete frequency distribution - barplot
  - Numeric - continuous frequency distribution - histogram
  
```{r}
library(ggplot2)
library(ggcorrplot)
library(ggbeeswarm)
```

```{r}
ggplot(data=df,aes(x=Gender))+geom_bar()
```
```{r fig.height=4,fig.width=8}

nums = c("CreditScore","Age","Tenure","Balance","NumOfProducts","EstimatedSalary")
cats = c("Geography","Gender","HasCrCard","IsActiveMember","Exited")

for (col in nums){
  print(ggplot(data=df,aes_string(x=col))+geom_histogram())
}

for (col in cats){
  print(ggplot(data=df,aes_string(x=col))+geom_bar())
}


```
Observation - 
  - approx 50 % of customers are from france, rest from spain and germany
  - approx 70% of customers have creditcards
  - approx 50% of customers are not active members
  - approx 20% of customers left the bank
  - Age has slightly right skewed distribution
  - Tenure has almost uniform distribution
  - there are profiles in context to balance, around 35% of customers have 0 balance, and rest have balance higher than 75k
  - most of the customers have either 1 or 2 products.
  - estimatedsalary has almost uniform distribution.


### Bivariate Analysis
  - numeric v/s numeric - scatterplot
  - numeric v/s categorical - continuous probability density distribution of numeric variable
  - categorical v/s categorical - discrete probability distribution of categorical variable

```{r fig.height=4,fig.width=8}
nums = c("CreditScore","Age","Tenure","Balance","NumOfProducts","EstimatedSalary")

df$Exited = factor(df$Exited)
for (col in nums){
  print(ggplot(data=df,aes_string(x=col,fill='Exited',alpha=0.1))+geom_density())
}
```

  categorical v/s categorical
```{r}
# Geography vs Exited
ggplot(data=df,aes(x=Geography))+geom_bar()+facet_grid(cols=vars(Exited),margins=TRUE)
```
```{r}
pivot = as.data.frame.matrix(table(df$Geography,df$Exited))
pivot[,2]/rowSums(pivot)
```
```{r}
cats = c("Geography","Gender","HasCrCard","IsActiveMember")
for (col in cats){
  print(col)
  pivot = as.data.frame.matrix(table(df[,col],df$Exited))
  print(pivot[,2]/rowSums(pivot))
  print(" ")
}
```
Multivariate Analysis
  - categorical v/s numeric vs numeric - scatterplot
  - categorical vs categorical vs numeric - swarmplot / boxplot / pointplot
```{r fig.width=12,fig.height=5}
# age v/s balance v/s exited
ggplot(data=df,aes(x=Age,y=Balance,col=Exited,alpha=0.1))+geom_point()
```
```{r fig.width=20,fig.height=8}
# age v/s geography v/s Exited
ggplot(data = df,aes(x=Geography,y=Age,col=Exited,aplha=0.5))+geom_beeswarm()
```
Observation -
Informative - Age, Balance, Numberofproducts, Geography, Gender, IsActiveMember

## Statistical Research
  - Correlation Analysis - linear patterns
  - ANOVA - f test
  - Chi Square test
```{r fig.width=15,fig.height=7}
# correlation Analysis
df$Exited = as.numeric(df$Exited)
corr = cor(df[,c(-2,-3)])
library(ggcorrplot)
ggcorrplot(corr,lab=TRUE)
```

### ANOVA - Analysis of Variance - f test

  - Numeric v/s cateogrical
  
- used to compare means of two or more groups
- for example - ANOVA can tell us whether mean of age of customers who left and mean of age of customers who did NOT leave are similar or different.

  - Null Hypothesis: means of two groups are similar
  - Alternate Hypothesis: means of two groups are not similar
  
When you apply ANOVA you get - pvalue
  
  For Business confidence (confidence Interval) of 95%
  CI = 95% = 0.95, alpha = 1 - CI = 1 - 0.95 = 0.05
  
    - if pvalue > alpha = We fail to reject null hypothesis - feature is NOT important
    - if pvalue < alpha = we succesfully reject null hypothesis - feature is important
    
```{r}
df$Exited = as.numeric(df$Exited)
nums = c("CreditScore","Age","Tenure","Balance","NumOfProducts","EstimatedSalary")
df_nums = df[,nums]
df_nums$Exited = df$Exited
anova_output = aov(Exited~.,data=df_nums)
# anova_op = aov(Exited~CreditScore+Age+Tenure+Balance+NumOfProducts+EstimatedSalary)
summary(anova_output)

```

### Chi Square Test
  
  - categorical v/s categorical
  - used to compare the distribution of a categorical variable in two or more groups
  - for example it can tell us, ratio of leaving vs not leaving is same for male vs female or not
  
  - Null Hypothesis: distribution of the categorical variable is similar in two or more groups
  - Alternate Hypothesis: distribution of the categorical variable is NOT similar in two or more groups
  
When you apply ANOVA you get - pvalue
  
  For Business confidence (confidence Interval) of 95%
  CI = 95% = 0.95, alpha = 1 - CI = 1 - 0.95 = 0.05
  
    - if pvalue > alpha = We fail to reject null hypothesis - feature is NOT important
    - if pvalue < alpha = we succesfully reject null hypothesis - feature is important
  
```{r}
cats = c("Geography","Gender","HasCrCard","IsActiveMember")
for (col in cats){
  print(col)
  print(chisq.test(x=df[,col],y=df$Exited))
}
```

  
  
  
  
  
  
  



